---
title: "U.S Births Data (1994 - 2014)"
author: "Claire Mondong"
date: "6/23/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Description

The data set that I am using the birth data in the United States between years 1994 - 2014. The data for years between 1994 - 2003 was provided by Centers for Disease Control and Prevention's National Center for Health Statistics while the data for years 2000 - 2014 was provided by the Social Security Administration.

5 columns of data were collected which includes the date, month and year, day of the week as well as the number of births from the years 1994 - 2014. The data records the number of births every single day between those years.The data were saved in a .csv file. It does not explicitly mention how data were collected but I am thinking through the number of birth certificates produced or through collecting data from hospitals throughout the United States.

The data regarding births was used to potentially prove that some people are too superstitious to have a baby on Friday the 13th. I am thinking that due to the negative connotations related to Friday the 13th, there are less families who want to give birth on that day.

## Reading the data into R

In order to read the file into R, 

I copy pasted the raw text file link from GitHub and inserted it into the R code using the function below and read the data into R as `birthdata` and `birthdata2` since the data were in 2 different files. 

```{r}
download.file(url = "https://raw.githubusercontent.com/fivethirtyeight/data/master/births/US_births_1994-2003_CDC_NCHS.csv", destfile = "birthdata.csv")
birthdata<-read.csv("birthdata.csv", stringsAsFactors = FALSE)
download.file(url = "https://raw.githubusercontent.com/fivethirtyeight/data/master/births/US_births_2000-2014_SSA.csv", destfile = "birthdata2.csv")
birthdata2<-read.csv("birthdata2.csv", stringsAsFactors = FALSE)
```

To test whether the data was imported to R, I entered the codes below to see the first 10 rows of each data file. 
```{r}
head(birthdata,10)
head(birthdata2,10)
```

Next, I tried renaming the column "date_of_month" to date to simplify or shorten the words for the column using the codes below.
```{r}
colnames(birthdata)[colnames(birthdata) =='date_of_month'] <- 'date'
colnames(birthdata2)[colnames(birthdata2) =='date_of_month'] <- 'date'
```

## Characteristics of the data
Through running the codes two codes below, the data had 5 columns which were "year", "month", "date_of_month", "day_of_week" and "births". Column 1 to 4 records the day, date, month and year while column 5 calculates the number of babies born in a day.
```{r}
colnames(birthdata)
colnames(birthdata2)
```

Using the codes below, the data for years 1994 - 2003 has 3652 rows while the data for years 2000 - 2014 has 5479 rows. 

```{r}
nrow(birthdata)
nrow(birthdata2)
```


The birth dataset for years between 1994 - 2003 dataset has `r ncol(birthdata)` columns and `r nrow(genderdata)` rows. 

The birth dataset for years between 2000 - 2014 dataset has `r ncol(birthdata2)` columns and `r nrow(genderdata2)` rows. 

The code below will generate the table of each column in the data set and a brief description of it.

```{r}
text_tbl <- data.frame(
  Names = c("year", "month", "date", "day_of_week", "births"),
  Description = c(
    "The year data was recorded",
    "The month data was recorded", 
    "The date of the month data was recorded",
    "The day of the week of the date recorded",
    "Number of births in that day"),
```

To check how the table appears, I used the code text_tbl. 

## Summary Statistics 

First, I checked whether the datasets had missing values using the code "sum(is.na(birthdata))" and "sum(is.na(birthdata2))". The code shows a result of 0, meaning that there are no missing values in the datasets. Then, using the codes summary(birthdata) and summary(birthdata2), I was able to obtain summaries of the columns in the datasets. 

```{r}
sum(is.na(birthdata)),

sum(is.na(birthdata2)),

summary(birthdata),

summary(birthdata2),
```

Lastly, using the code print(summary(birthdata)) and print(summary(birthdata2)) would show the summary statistics. 

```{r}
print(summary(birthdata)),

print(summary(birthdata2))
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
